### 2016-04-14 manuscript_v0.Rmd
again

---
title: "Testing new Pauly Cessational Growth Function"
author: "Derek H. Ogle"
date: "April 9, 2016"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
  html_document:
    fig_height: 3.5
    fig_width: 4.5
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738423)  # for reproducibility
library(FSA)
library(FSAsim)

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("vbMain")
eqncaps("vbMain2")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers2")
```

```{r captions, include=FALSE}

```

```{r maketprFun, echo=FALSE}
################################################################################
## internal function to compute t-prime
################################################################################
iCalc_tpr <- function(t,WP,NGT) {
  ## No-growth period starts at WP-NGT/2.  Use this to shift each age such that
  ##    first NGT would correspond to a value of 0
  tshift <- WP-NGT/2
  tmp.t <- t-tshift
  ## Find fraction of year on this new time scale for each adjusted age
  ##    by substracting whole number age (e.g, 2.79-2)
  tmp.t2 <- tmp.t-floor(tmp.t)
  ## Adjust this fraction for no growth ... with the time adjustment no-growth
  ##    is from 0 to NGT in each "year" ... tprime will not increase in NGT so
  ##    replace all of these fractions with zeroes.  Above NGT, the tprime must
  ##    have the NGT removed from it (e.g., if NGT=0.4, then 0.79-0.4 is the
  ##    fraction of the year for which growth occurred.)
  for (i in 1:length(tmp.t2)) {
    if (tmp.t2[i]<=NGT) tmp.t2[i] <- 0
    else tmp.t2[i] <- tmp.t2[i]-NGT
  }
  ## Now get the number of completed growth periods (i.e., floor (tmp.t)) times
  ##    the length of a growth period (1-NGT) to get the total growth periods
  ##    completed prior to that year and then add on the fractional piece
  tmp.t <- floor(tmp.t)*(1-NGT)+tmp.t2
  ## Shift back to get tprime (why is NGT needed here)
  tpr <- tmp.t+tshift+NGT
}
```

```{r makePaulyFun, echo=FALSE}
################################################################################
## Main Function
##   Linf, t0 as usual
##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
##   WP = "Winter Period" (middle point of no-growth period)
##      = ts+0.5 (ts from Pauly et al.)
##   NGT = "No Growth Time" = "fraction of a year where no growth occurs"
##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
################################################################################

paulySC <- function(t,Linf,Kpr=NULL,t0=NULL,WP=NULL,NGT=NULL) {
  if (length(Linf)==5) { Kpr <- Linf[[2]]; t0 <- Linf[[3]]
                         WP <- Linf[[4]]; NGT <- Linf[[5]]
                         Linf <- Linf[[1]] }
  tpr <- iCalc_tpr(t,WP,NGT)
  q <- Kpr*(tpr-t0) +
    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-WP-0.5)) -
    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-WP-0.5))
  Linf*(1-exp(-q))
}
```

# Introduction

The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).  The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$

with

$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.

Many animals exhibits seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).  `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sin function, to model these seasonal oscillations in growth.  The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses

$$
\begin{aligned}
q & =K(t-t_{0}) \\
  & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`} \\
  & -\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))
\end{aligned}  
$$

where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.55; Linf <- 30; K <- 0.3; t0 <- -0.1
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\

Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].  To address this issue, @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period with a smooth transition of the modeled mean length-at-age into and out of the no-growth period.  Specifically, their modification is

\

$$
\begin{aligned}
q & =K^{'}(t^{'}-t_{0}) \\
  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`} \\
  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right)
\end{aligned}  
$$

\

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` because the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` due to $t^{'}$&lt;$t$ for any $t$.  We further note that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.

@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth), and replacing $K$ with $K^{'}$.  Their modification can be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).  The growth trajectory is then separated at the $WP$ and a horizontal segment that is $NGT$ units long is inserted.  This forms a growth trajectory that smoothly transitions into and out of a no-growth period (`r figcaps("PaulyExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
WP <- 0.55; NGT <- 0.3
Linf <- 30; Kpr <- 0.4; t0 <- -0.1
Pcf <- c(Linf,Kpr,t0,WP,NGT)
## Get growth trajectory
t <- seq(-0.1,3.3,length.out=499)
PL <- paulySC(t,Linf=Pcf)
## Make schematic plot
plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
     ylab="Length",xlab="Age (years)")
axis(1,0:3)
## Put on NGT polygons
# find WPs and Lengths at WPs
WPs <- 0:2+WP
LatWPs <- paulySC(WPs,Linf=Pcf)
# find start and end of NGT
NGTs <- WPs-NGT/2
NGTe <- WPs+NGT/2
# place polygons
for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
                                  col=col2rgbt("black",1/20),border=NA)
## Put on the growth trajectory
lines(PL~t,lwd=2)
## Place an open point at the WP
points(WPs,LatWPs,pch=3,col="gray50")
## Add some important values to the age axis
axis(1,NGTs,tcl=-0.2)
axis(1,NGTe,tcl=-0.2)
## Add the t-prime labels
axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
axis(1,at=0:3,labels=c(NA,0.7,1.4,2.1),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTs,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTe,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=3.3,labels="<= t",tick=FALSE)
axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)

figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".  Each no-growth period is marked by a light gray polygon and each $WP$ is marked by the gray plus symbol.  The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
```

`r figcaps("PaulyExPlot")`

\

@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.  The diskette is difficult (at best) to obtain and the source code is no longer available (D. Pauly, pers. comm.).  @Paulyetal1992 did describe the operations performed by their program, but their is no description of how $t^{'}$ was operationalized.  This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.  In other words, the values for $t^{'}$ will change with each iteration in the non-linear model-fitting process.

The objectives of this note are to (i) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (ii) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.

# The Modified Model

The $WP$ value is equal to $t_{s}+0.5$ because of the shape of the sin function used to model the growth oscillations.  By simple substitution, `r eqncaps("PaulyMod",display="cite")` is modified to include $WP$ rather than $t_{s}$.

\

$$
\begin{aligned}
q & =K^{'}(t^{'}-t_{0}) \\
  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-WP-0.5)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-WP-0.5)\right)
\end{aligned}  
$$

\

Note that in `r eqncaps("PaulyMod2",display="cite")` $WP$ is the center of the no-growth period (`r figcaps("PaulyEx",display="cite")`).

# Methods
## Simulated Data

```{r echo=FALSE}
WP <- 0.8
NGT <- 0.5
gper <- c(WP+NGT/2-1,WP-NGT/2)*365
## Set some growth parameters
Linf <- 30; K <- 0.3; t0 <- 0
## Actually generate the data (with little variability)
d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.01,SD=0.01,paramCV=0.01,
                growth.per=gper,sample.per=c(1,365),lendigs=1)
```

## R Implementation

# Model Fitting
## Simulated Data

```{r echo=FALSE, message=FALSE}
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
## Try fitting the PaulySC model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.8,NGT=0.1)
fitP1 <- nls(lenCap~paulySC(ageFracY,Linf,K,t0,WP,NGT),data=d1,
             start=sv,lower=lwrbnd,upper=uprbnd,algorithm="port",control=list(maxiter=100))
## Make a diagnostic plot
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE)
cfs <- coef(fitP1)
WPs <- 1:5+cfs[["WP"]]
LatWPs <- paulySC(WPs,Linf=cfs)
segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
SNGTs <- WPs-cfs[["NGT"]]/2
ENGTs <- WPs+cfs[["NGT"]]/2
segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")
legend("topleft",paste(c("WP","NGT"),c(WP,NGT),sep="="),
       bty="n",inset=-0.02,text.col=c("red","blue"))

cbind(Est=coef(fitP1),confint(fitP1))
```
```{r echo=FALSE}
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5),xlim=c(1.7,1.9),ylim=c(13.425,13.43))
curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE,n=1001)
segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")

```

```{r echo=FALSE, message=FALSE}
## Try fitting the Somers model
vbSO <- vbFuns("Somers2")
fitSO1 <- nls(lenCap~vbSO(ageFracY,Linf,K,t0,C,WP),data=d1,
               start=list(Linf=32,K=0.32,t0=0,C=0.8,WP=0.8),
               lower=c(Linf=0,K=0,t0=-Inf,C=-1,WP=0),
               upper=c(Linf=Inf,K=Inf,t0=Inf,C=2,WP=1),
               algorithm="port",control=list(maxiter=100),trace=FALSE)
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
curve(vbSO(x,coef(fitSO1)),from=1,to=6,col="red",lwd=2,add=TRUE)
cbind(coef(fitSO1),confint(fitSO1))
```

## Real Data

## Appendix

```{r}
<<maketprFun>>
```

```{r}
<<makePaulyFun>>
```

## References