### 2016-05-22 zzzOld/doc/manuscript_v0.Rmd
Moved to Word, Reorganized file strux, added mosquitofish example

---
title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
author: "Derek H. Ogle"
date: "April 29, 2016"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
  html_document:
    fig_height: 3.5
    fig_width: 4.5
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738422)  # for reproducibility
library(FSA)
library(FSAsim)
library(latex2exp)
library(knitr)

source("iCalc_tpr.R")
source("VBSCGF.R")

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("vbMain")
eqncaps("vbMain2")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers2")
```

# Introduction

The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).  The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$

with

$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.

Many animals exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).  `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.  The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses

$$
\begin{aligned}
q= & K(t-t_{0}) \\
   & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`}
\end{aligned}  
$$

where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.55; Linf <- 30; K <- 0.3; t0 <- -0.1
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Mean Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\

Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].  To address this unrealistic decrease in mean length, @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.  Specifically, their modification is

\

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right) \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`}
\end{aligned}  
$$

\

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.  To eliminate confusion, he suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.

@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates in this shorter period).  Their modification may be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).  The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.  This forms a growth trajectory that smoothly transitions into and out of the no-growth periods (`r figcaps("PaulyExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
WP <- 0.55; NGT <- 0.3
Linf <- 30; K <- 0.3; Kpr <- K/(1-NGT); t0 <- 0
Pcf <- c(Linf,Kpr,t0,WP,NGT)
## Get growth trajectory
t <- seq(-0.1,3.3,length.out=499)
PL <- VBSCGF(t,Linf=Pcf)
## Calculate WPs and start and end of NGT
# find WPs and Lengths at WPs
WPs <- 0:2+WP
LatWPs <- VBSCGF(WPs,Linf=Pcf)
# find start and end of NGT
NGTs <- WPs-NGT/2
NGTe <- WPs+NGT/2
## Make schematic plot
plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
     ylab="Length",xlab="Age (years)")
## Put on NGT polygons
for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
                                  col=col2rgbt("black",1/20),border=NA)
## Put on the growth trajectory
lines(PL~t,lwd=2)
## Place a point at the WP
points(WPs,LatWPs,pch=3,col="gray50")
## Add some important values to the age axis
axis(1,0:3)
axis(1,NGTs,tcl=-0.2)
axis(1,NGTe,tcl=-0.2)
## Add the t-prime labels
# makes inside ticks
axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
axis(1,at=1:3,labels=iCalc_tpr(1:3,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTs,labels=iCalc_tpr(NGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTe,labels=iCalc_tpr(NGTe,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=3.3,labels="<= t",tick=FALSE)
axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)

figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",formatC(Kpr,format="f",digits=4),", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".  The no-growth periods are marked with gray bars and each $WP$ is marked with a gray plus symbol.  The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
```

`r figcaps("PaulyExPlot")`

\

@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.  The diskette is difficult (at best) to obtain and the source code is no longer available (D. Pauly, pers. comm.).  @Paulyetal1992 did describe the operations performed by their program, but there is no description of how $t^{'}$ was operationalized.  This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.  In other words, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.

Therefore, the objectives of this note are to (i) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (ii) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.

# The Modified Model

In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period of one year.  The growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$.  Thus, the start of the no-growth period ($SNG$) is $t_{s}+\frac{1-NGT}{2}$.  The center of the no-growth period is then $SNG+\frac{NGT}{2}$ or $t_{s}+\frac{1}{2}$.  Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period for `r eqncaps("PaulyMod",display="cite")`.

By simple substitution, `r eqncaps("PaulyMod",display="cite")` may be modified to include $WP$ rather than $t_{s}$.

\

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t^{'}-WP-\frac{1}{2}\right)\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t_{0}-WP-\frac{1}{2}\right)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`}
\end{aligned}  
$$

\

As noted by @Paulyetal1992 the calculation of $t^{'}$ depends on the observed age ($t$) and the cumulative no-growth time prior to $t$.  In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.  Here, the position of the no-growth period is defined by $SNG$.  However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$ it can also be defined relative to these parameters.

The following algorithm is used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).

1. Shift the age ($t$) by subtracting the start of the no-growth period ($SNG$) from $t$, such that a whole number age represents the start of a no-growth period.  For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
1. Subtract the whole number age from the shifted age such that the remaining decimal represents the fraction of a shifted year.  For example, a 0 will result if the shifted age is 2.0 and a 0.5 will result if the shifted age is 2.5.
1. Substract the $NGT$ from the value from the previous step.
1. If the value from the previous step is negative, then the age is within the no-growth period and the negative value should be replaced with a zero.  Otherwise, the positive value represents the fraction of the growth period completed.
1. Add the value from the previous step to the total growth time completed (i.e., the product of the number of growth periods completed and the length of the growth period ($1-NGT$)).
1. Compute $t{'}$ by adding back the $SNG$ that was subtracted in Step 1.

Further examples of $t^{'}$ values relative to $t$ values are shown in `r figcaps("PaulyExPlot",display="cite")`.  This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.  With this, `r eqncaps("paulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.

# Fitting the Modified Model
Simple simulated data were used to demonstrate the fitting of `r eqncaps("paulyMod2",display="cite")`.  The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("paulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.  The `nls()` function from R was then used to estimate parameter values for `r eqncaps("paulyMod2",display="cite")` from the simulated data.  The "port" algorithm in `nls()` was used as $L_{\infty}$ and $K^{'}$ were constrained to be positive and $WP$ and $NGT$ were constrained to be between 0 and 1.  One would expect the parameter estimates to be very close to the values used to create the simulated data as `r eqncaps("paulyMod2",display="cite")` was used to both create and fit the data.

One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulyExData1",display="cite")`.  In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence interval for each parameter).

```{r echo=FALSE, results='hide'}
## Set some common parameters
Linf <- 30; K <- 0.3; t0 <- -0.1
## Set bounds for model-fitting
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
```
```{r echo=FALSE, results="hide", message=FALSE}
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
sigma <- 0.3
ages <- runif(200,min=0,max=5)
lens <- VBSCGF(ages,Linf,K,t0,WP,NGT)+rnorm(200,0,sigma)
dB1 <- data.frame(ages,lens)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitB1 <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=dB1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfs <- cbind(Est=coef(fitB1),confint(fitB1))
## A plot
# Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(lens~ages,data=dB1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(0,5),ylim=c(0,18))
curve(VBSCGF(x,cfs[,"Est"]),from=0,to=6,lwd=2,add=TRUE)
WPs <- 0:5+cfs["WP","Est"]
LatWPs <- VBSCGF(WPs,Linf=cfs[,"Est"])
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs["NGT","Est"]/2
ENGTs <- WPs+cfs["NGT","Est"]/2
arrows(SNGTs,LatWPs+1.2,ENGTs,LatWPs+1.2,lwd=2,length=0.025,angle=90,code=3)

legend("topleft",c(TeX("\\mathbf{Set Values}"),
                   TeX(paste("$L_{\\infty}$",Linf,sep=" = ")),
                   paste("K'",K,sep=" = "),
                   TeX(paste("$t_{0}",t0,sep=" = ")),
                   paste("WP",WP,sep=" = "),
                   paste("NGT",NGT,sep=" = ")),
        bty="n",inset=-0.02,cex=0.8)
legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
                       TeX(paste0("$L_{\\infty}$"," = ",
                                  formatC(cfs["Linf","Est"],format="f",digits=1),
                                  " (",formatC(cfs["Linf","2.5%"],format="f",digits=1),
                                  ", ",formatC(cfs["Linf","97.5%"],format="f",digits=1),")")),
                   paste0("K' = ",formatC(cfs["K","Est"],format="f",digits=2),
                                  " (",formatC(cfs["K","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["K","97.5%"],format="f",digits=2),")"),
                   TeX(paste0("$t_{0}$"," = ",
                                  formatC(cfs["t0","Est"],format="f",digits=2),
                                  " (",formatC(cfs["t0","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["t0","97.5%"],format="f",digits=2),")")),
                   paste0("WP = ",formatC(cfs["WP","Est"],format="f",digits=2),
                                  " (",formatC(cfs["WP","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["WP","97.5%"],format="f",digits=2),")"),
                   paste0("NGT = ",formatC(cfs["NGT","Est"],format="f",digits=2),
                                  " (",formatC(cfs["NGT","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["NGT","97.5%"],format="f",digits=2),")")),
        bty="n",cex=0.8)

figcaps("PaulyExData1",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".  Paramter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.  Each WP is shown by the arrow and each NGT is shown by the horizontal interval bar centered on the WP arrow."))
```

`r figcaps("PaulyExData1")`

\

```{r echo=FALSE, message=FALSE}
WPs <- rep(c(0.8,0.2),each=4)
NGTs <- rep(rep(c(0.5,0.2),each=2),2)
sigmas <- rep(c(0.3,0.5),4)
ages <- runif(200,min=0,max=5)
sv <- list(Linf=32,K=0.32,t0=0,WP=0.5,NGT=0.5)

## Start with results from above
res <- matrix(nrow=8,ncol=18)
res[1,] <- c(WPs[1],NGTs[1],sigmas[1],round(cfs["WP",],2),round(cfs["NGT",],2),round(cfs["Linf",],1),round(cfs["K",],2),round(cfs["t0",],2))
for (i in 2:length(WPs)) {
  tmp <- data.frame(ages,lens=VBSCGF(ages,Linf,K,t0,WPs[i],NGTs[i])+rnorm(200,0,sigmas[i]))
  tmpfit <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=tmp,
                start=sv,lower=lwrbnd,upper=uprbnd,
                algorithm="port",control=list(maxiter=100))
  tmpcfs <- cbind(Est=coef(tmpfit),confint(tmpfit))
  res[i,] <- c(WPs[i],NGTs[i],sigmas[i],round(tmpcfs["WP",],2),round(tmpcfs["NGT",],2),round(tmpcfs["Linf",],1),round(tmpcfs["K",],2),round(tmpcfs["t0",],2))
}
colnames(res) <- c("WPset","NGTset","sigmaset","WPest","WPlci","WPuci","NGTest","NGTlci","NGTuci","Linfest","Linflci","Linfuci","Kest","Klci","Kuci","t0est","t0lci","t0uci")
res
```

```{r echo=FALSE, results='hide',eval=FALSE}
## Set some common parameters
Linf <- 30; K <- 0.3; t0 <- -0.1
## Set bounds for model-fitting
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
```

```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
gper <- c(WP+NGT/2-1,WP-NGT/2)*365
## Actually generate the data (with very little variability)
d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.001,SD=0.001,paramCV=0.001,
                growth.per=gper,sample.per=c(1,365),lendigs=1)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
( cfs <- coef(fitP1) )
## A plot
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
WPs <- 1:5+cfs[["WP"]]
LatWPs <- VBSCGF(WPs,Linf=cfs)
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs[["NGT"]]/2
ENGTs <- WPs+cfs[["NGT"]]/2
arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)

tmp <- c("Linf=","K=","t0=","WP=","NGT=")
legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
       bty="n",inset=-0.02,cex=0.8)
cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
```

```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
gper <- c(WP+NGT/2-1,WP-NGT/2)*365
## Actually generate the data (with more variability)
d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
                growth.per=gper,sample.per=c(1,365),lendigs=1)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
( cfs <- coef(fitP1) )
## A plot
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
WPs <- 1:5+cfs[["WP"]]
LatWPs <- VBSCGF(WPs,Linf=cfs)
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs[["NGT"]]/2
ENGTs <- WPs+cfs[["NGT"]]/2
arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)

tmp <- c("Linf=","K=","t0=","WP=","NGT=")
legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
       bty="n",inset=-0.02,cex=0.8)
cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
```

```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
gper <- c(WP+NGT/2-1,WP-NGT/2)*365
## Actually generate the data (with more variability)
d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
                growth.per=gper,sample.per=c(1,365),lendigs=1)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitP1 <- nls(lenCap~VBSCGF(ageFracY-0.3,Linf,K,t0,WP,NGT),data=d1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
( cfs <- coef(fitP1) )
## A plot
plot(lenCap~I(ageFracY-0.3),data=d1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
WPs <- 1:5+cfs[["WP"]]
LatWPs <- VBSCGF(WPs,Linf=cfs)
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs[["NGT"]]/2
ENGTs <- WPs+cfs[["NGT"]]/2
arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)

tmp <- c("Linf=","K=","t0=","WP=","NGT=")
legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP-0.3,NGT)),
       bty="n",inset=-0.02,cex=0.8)
cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
```

## Real Data

# Discussion

# Appendices
## Appendix 1

```{r echo=FALSE, comment=""}
cat(readLines("iCalc_tpr.R"),sep="\n")
```

## Appendix 2

```{r echo=FALSE, comment=""}
cat(readLines("VBSCGF.R"),sep="\n")
```

# References